---
title: "Montagu API from R"
author: "VIMC Technical Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Montagu API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this vignette, we describe a typical example of building
a model in R, which uses the demography and templates available
through the Montagu API, and submited burden estimates to it. We
will proceed very naively and cover all of the important
functionality on the way.

### Identify the Montagu Server

Begin by identifying the Montagu server to connect to. You must
have an account on this server; typically this will be our live
production server, and you will have received login information
on joining VIMC. Identify the server as follows:-

```{r}
montagu::montagu_server_global_default_set(
  montagu::montagu_server("production", "montagu.vaccineimpact.org"))
```

The next time you query the server, you will be asked to
login with a username and password.

### Basic Information

For many of the montagu api functions, you will need to provide your
modelling group id, and your disease id. These will have been communicated on
joining VIMC, but can be looked up with simple function calls like this:-

```{r}
montagu::montagu_diseases()
```

TO-DO - i2557

```r
montagu::montagu_modelling_groups()
```

### Touchstones

A touchstone is an identifier that links:

* A request from VIMC to the modelling groups to provide burden estimates
* The input data (coverage and demography) provided to calculate those estimates
* The estimates submitted by the groups using the input data.

Touchstone ids are referred to as a basename and a version together. Errata
to the input data may be addressed with a new touchstone version. Generally,
modelling groups will know which touchstone they are required to run models
against. To lookup all the currently open touchstones for a group:-

```{r}
montagu::montagu_touchstone_versions("IC-Garske", require_open = TRUE)
```

If we knew the basic touchstone name (without version) that we were interested in,
then we could also have asked:
```{r}
montagu::montagu_touchstone_versions("IC-Garske", "201710gavi", require_open = TRUE)
```

### Scenarios

Within a touchstone, we may have various scenarios, describing different
vaccination conditions. These vary across diseases due to the different types
of vaccination approaches that have been employed. We look our the scenarios
we are required to model for a touchstone with:

```{r}
montagu::montagu_scenarios("IC-Garske", "201710gavi-5")
```

We can then look up the status a particular scenario within this touchstone:

```{r}
montagu::montagu_scenario_status("IC-Garske", "201710gavi-5", "yf-routine-gavi")
```

and if that status were not valid, then we can query any problems with:
```{r}
montagu::montagu_scenario_problems("IC-Garske", "201710gavi-5", "yf-routine-gavi")

```

### Expectations - years, ages, countries and outcomes

For a given touchstone, modelling groups are required to produce burden estimates
stratified by calendar year, and by age. The expectations about time and age
will be consistent across scenarios. We can look up these expectations like so:

```{r}
montagu::montagu_expectations("IC-Garske", "201710gavi-5")
```

Some modelling groups model more than one disease, in which case a row appears
for each disease. And additionally, if there is some reason to have different
expectations for different scenarios, then multiple lines for the same disease
may appear, with a difference in the description column indicating the scenario.
For example:
```{r}
montagu::montagu_expectations("CDA-Razavi", "201710gavi-5")
```

If we already knew the expectation id, we could look up
a single expectation with

```{r}
montagu::montagu_expectation("IC-Garske", "201710gavi-5", 30)
```

The countries for which estimates are required can vary between disease, and
between scenarios. To retrieve the list of countries that are required for a
particular expectation:-

```{r}
countries <- montagu::montagu_expectation_countries("IC-Garske", "201710gavi-5", 30)
head(countries)
```

To see which outcomes are required for an expectation:
```{r}
montagu::montagu_expectation_outcomes("IC-Garske", "201710gavi-5", 30)
```

And to confirm which scnearios an expectation applies to (which should align
with the description text for the expectation):
```{r}
montagu::montagu_expectation_applicable_scenarios("IC-Garske", "201710gavi-5", 30)
montagu::montagu_expectation_applicable_scenarios("CDA-Razavi", "201710gavi-5", 2)
```

### Templates

The above will give all the information required to build your own data frame
for submitting the expected results. Alternatively, you can download
burden estimate templates with the years, countries and ages already populated:-

```{r}
df <- montagu::montagu_central_burden_estimate_template("IC-Garske", "201710gavi-5", 30)
head(df)
```
